---
title: "simulation"
author: "Azad Sadr"
date: '2022-02-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r get_utilities, echo=FALSE, include=FALSE}
source("utilities.R")
```

### Load data
```{r, echo=FALSE}
json_path = "/home/azad/Documents/thesis/SigPhylo/data/results/output5/output.json"
data <- load_data(json_path)
data
```

### Overall insights on :
k, $\lambda$, log-likelihood, BIC & Constructed Phylogeny
```{r, echo=FALSE}
plot_k_loglike(data)
plot_k_bic(data)
plot_lambda_loglike(data)
plot_lambda_bic(data)
heatmap(data)
plot_cosine(data, 0.95)
```

### Best Run info (log-likelihood)
```{r, echo=FALSE}
best_loglike_info <- best_loglike(data)
k_loglike <- best_loglike_info[["k"]]
lambda_loglike <- best_loglike_info[["lambda"]]

print(paste("Best K_Denovo :", k_loglike))
print(paste("Best Lambda :", lambda_loglike))
```

### Best Run info (BIC)
```{r, echo=FALSE}
best_bic_info <- best_bic(data)
k_bic <- best_bic_info[["k"]]
lambda_bic <- best_bic_info[["lambda"]]

print(paste("Best K_Denovo :", k_bic))
print(paste("Best Lambda :", lambda_bic))
```

### Visualize Alpha for Best BIC
```{r, echo=FALSE}
alpha <- alpha_read_tibble(data, k_bic, lambda_bic) # it seems no need to this function
plot_alpha(data, k_bic)
```

### Visualize Fixed Beta
```{r, echo=FALSE}
cosmic_path <- "/home/azad/Documents/thesis/SigPhylo/cosmic/cosmic_catalogue.csv"
SBS5 <- beta_read_csv(cosmic_path)["SBS5", ]  # SBS5 from cosmic
plot_beta(SBS5)
```

```{r, echo=FALSE}
# expected beta
beta_expected_path <- "/home/azad/Documents/thesis/SigPhylo/cosmic/cosmic_catalogue.csv"
beta_expected <- beta_read_csv(beta_expected_path)

# expected alpha
alpha_expected_path <- "/home/azad/Documents/thesis/SigPhylo/cosmic/cosmic_catalogue.csv"
alpha_expected <- beta_read_csv(alpha_expected_path)
```


```{r, echo=FALSE}
# label inferred beta (lowest BIC)
beta_denovo <- beta_read_tibble(data, k_bic, lambda_bic)
beta_denovo_labeled <- beta_labeling(beta_expected, beta_denovo)
```

```{r, echo=FALSE}


```


```{r,echo=FALSE}
# Inferred vs Expected Beta
# SBS1 -------------------------------------------------------------------------
glist_SBS1 <- list()
glist_SBS1[[1]] <- plot_beta(SBS1)
glist_SBS1[[2]] <- plot_beta(beta_denovo_labeled["SBS1", ])

a1 <- as.numeric(SBS1)  # numeric vector of expected SBS1
b1 <- as.numeric(beta_denovo_labeled["SBS1", ])  # numeric vector of inferred SBS1
cosine_SBS1 <- cosine_sim(a1, b1)

gplot_SBS1 <- ggarrange(plotlist = glist_SBS1, 
                        ncol = 1, 
                        #nrow = nrow, 
                        common.legend = TRUE, 
                        legend = "bottom")

# Unknown ----------------------------------------------------------------------
glist_Unknown <- list()
glist_Unknown[[1]] <- plot_beta(beta_expected["Unknown", ])
glist_Unknown[[2]] <- plot_beta(beta_denovo_labeled["Unknown", ])

a2 <- as.numeric(beta_expected["Unknown", ])  # numeric vector of expected Unknown
b2 <- as.numeric(beta_denovo_labeled["Unknown", ]) # numeric vector of inferred Unknown
cosine_Unknown <- cosine_sim(a2, b2)

gplot_Unknown <- ggarrange(plotlist = glist_Unknown, 
                           ncol = 1, 
                           #nrow = nrow, 
                           common.legend = TRUE, 
                           legend = "bottom")
```

### Visualize Inferred vs. Expected SBS1 Signature
first plot represents expected signature and second plot represents inferred signature.
```{r, echo=FALSE}
print(paste("cosine similarity between two signatures :", cosine_SBS1))
gplot_SBS1
```

### Visualize Inferred vs. Expected Unknown Signature
first plot represents expected signature and second plot represents inferred signature.
```{r, echo=FALSE}
print(paste("cosine similarity between two signatures :", cosine_Unknown))
gplot_Unknown
```


